---
title: "lab_9"
author: "Alyssa Kibbe"
date: "11/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(palmerpenguins)
library(GGally)
library(stargazer)
library(here)
library(broom)
library(janitor)
```
# Part 1

Exploring the data
```{r}
penguins %>% 
  ggpairs(aes(color = species))
```

Restricted exploring
```{r}
penguins %>% 
  select(species, bill_length_mm:body_mass_g) %>% 
  ggpairs(aes(color = species))
```
creating linear models 
1. penguin mass ~ flipper length + species
2. penguin mass ~ flipper length + species + sex
3. penguin mass ~ flipper length + species + sex + bill length
4. penguin mass ~ flipper length + species + sex + bill length + island

```{r}
lm1 <- lm(body_mass_g ~ flipper_length_mm + species, data = penguins)

lm2 <- lm(body_mass_g ~ flipper_length_mm + species + sex, data = penguins)

lm3 <- lm(body_mass_g ~ flipper_length_mm + species + sex + bill_length_mm, data = penguins)

lm4 <- lm(body_mass_g ~ flipper_length_mm + species + sex + bill_length_mm + island, data = penguins)
```

```{r}
plot(lm1) # no concerns
plot(lm2) # no concerns
plot(lm3) # no concerns
plot(lm4) #no concerns
```
AIC
```{r}
AIC(lm1) # 5031.523
AIC(lm2) # 4740.774
AIC(lm3) # 4733.574
AIC(lm4) # 4736.979
```

stargazer package to show side by side results
```{r}
stargazer(lm1, lm3, type = "html")
```
omitting a variable
```{r}
ggplot(data = penguins, aes(x = flipper_length_mm, y = bill_depth_mm))+
  geom_point()+
  geom_smooth(method = "lm")

#include species variable
ggplot(data = penguins, aes(x = flipper_length_mm, y = bill_depth_mm, group = species)) +
  geom_point(aes(color = species))+
  geom_smooth(method = "lm")
```
# Part 2

```{r}
lizards <- read.csv(here("data", "lizards.csv"))

lizards %>% 
  count(spp) %>% 
  arrange(-n)

# UTST = Uta stansburiana (side-blotched lizard)
```

 subgroups
```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail %in% c("B", "W"))
```
 
 counts & proportions & making table
 
```{r}
tail_counts <- utst %>% 
  janitor::tabyl(site, tail)

# both percentages & counts
tail_proportions <- tail_counts %>% 
  adorn_percentages() %>% 
  janitor::adorn_pct_formatting(digits = 2) %>% 
  adorn_ns()
```
 
contingency table & chi-square
 -- format to chi-square table
```{r}
lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site")
```
run test
```{r}
lizard_x2 <- chisq.test(lizard_ct)
lizard_x2
lizard_tidy <- tidy(lizard_x2)
lizard_tidy
```




